Are kelp forests different? A comparison of kelp forests between contrasting study sites

### 2018-07-18 paper_draft.Rmd
reference formatting, more to come...

Kelps are a cold-water species of large brown algae, belonging to the order Laminariales (Kelly 2005). They are regarded as ecosystem engineers, providing a 3-dimensional habitat for a variety of fauna and flora. In addition, kelp ecosystems host a number of economically important species such as lobster, abalone and fish (Dayton 1985, Steneck et al. 2002). Various abiotic (e.g. light, wave exposure, sedimentation) and biotic (e.g. grazing) factors structure kelp forests (Devinny and Volse 1978, Dayton et al. 1992, Utter and Denny 1996, Toohey et al. 2004, Rothman et al. 2006, Sjøtun et al. 2006, Wernberg and Goldberg 2008, Pedersen et al. 2012). Variations in these factors ultimately affects the way kelp-associated benthos communities are structured (Dayton 1985, Guimaraes and Coutinho 1996, Wernberg and Goldberg 2008, Wernberg 2009).
Temperature is recognised as the main abiotic driver of marine species distributions (Dayton et al. 1992, Hiscock et al. 2004, Wernberg et al. 2011, 2013, 2016, Mabin et al. 2013). Given kelp’s preference for cold-temperate waters, ocean warming poses a serious threat to these ecosystems (Wernberg et al. 2010, 2011, Vergés et al. 2016). In the absence of kelp, other types of systems may persist, for example algal turf, coralline reefs or urchin barrens (Vergés et al. 2016). These other systems support a different array of fauna and flora and are considered less diverse and less productive than kelp forests (Dayton 1985, Connell et al. 2008, Byrnes et al. 2011). In future climate scenarios, species distributions are expected to shift poleward in response to changing ocean temperatures. Given kelps are cold-water species, changes in ocean temperatures, particularly warming, could have detrimental effects on their growth, survival, reproduction and recruitment.

Other abiotic drivers such as light (Dayton et al. 1984, Bolton and Levitt 1985, Dieck 1993, Sjøtun et al. 2006, Wernberg and Goldberg 2008), sedimentation (Duggins et al. 1990, Vadas et al. 1992, Schiel and Foster 2006, Schiel et al. 2006) and water motion (Hurd 2000, Kawamata 2001, Mudge and Scrosati 2003, Scrosati and Mudge 2004, Wernberg and Goldberg 2008, Pedersen et al. 2012) have been shown to be important in the settlement, growth, and survival of kelp and biotic variables (e.g. see Black 1976, Kennelly 1983, Dayton et al. 1984, Harris et al. 1984, Carpenter 1990, Sivertsen 2006, Harley et al. 2012) play an equally important role in structuring kelp communities, through alteration in top-down and bottom-up control mechanisms (Breen and Mann 1976, Duggins 1980, Harris et al. 1984, Estes and Duggins 1995, Shears and Babcock 2002, Arkema et al. 2009, Blamey and Branch 2012).

@@

### Statisical analyses

@@

### Abiotic parameters

@@

flora_df <- read.csv(file = "data/flora.csv", check.names = FALSE)
fauna_df <- read.csv(file = "data/fauna.csv", check.names = FALSE)
fauna_df <- fauna_df[ ,c(1:27)]
benthic <- read.csv(file = "data/benthic.csv", check.names = FALSE)
cover_df <- read.csv(file = "data/cover.csv", check.names = FALSE)
kelp_df <- read.csv(file = "data/kelp.csv", check.names = FALSE)

@@

### Diversity

Diveristy of flora was...

@@

plot_flora <- ggplot(data = flora_lng, aes(x = species, y= biomass, fill = site)) + 
              geom_bar(stat = "identity", show.legend = FALSE, position=position_dodge()) + 
              theme(axis.text.x=element_text(angle = 90, hjust = 1, vjust = 0.3)) +
              theme(panel.background = element_rect(fill = 'white', colour = 'white')) +
              ylab("") + 
              xlab("") +
              coord_flip()
```

Diveristy of fauna was.....

```{r}

@@

```{r biomass barplot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap='Comparison of raw biomass species data between sites. Graph A represents flora and graph B represents fauna.'}
ggarrange(plot_flora, plot_fauna, ncol = 2, nrow = 1, labels = "AUTO", legend = "bottom", common.legend = TRUE)
```

>>>>

Kelps are a cold-water species of large brown algae, belonging to the order Laminariales [@Kelly 2005]. They are regarded as ecosystem engineers, providing a 3-dimensional habitat for a variety of fauna and flora. In addition, kelp ecosystems host a number of economically important species such as lobster, abalone and fish [@Dayton1985; @Steneck2002]. Various abiotic (e.g. light, wave exposure, sedimentation) and biotic (e.g. grazing) factors structure kelp forests [@Devinny1978; @Dayton1992; @Utter1996; @Toohey2004; @Rothman2006; @Sjøtun2006; @Wernberg2008; @Pedersen2012]. Variations in these factors ultimately affects the way kelp-associated benthos communities are structured [@Dayton1985; @Guimaraes1996; @Wernberg2008; @Wernberg 2009]
.
Temperature is recognised as the main abiotic driver of marine species distributions [@Dayton1992; @Hiscock2004; @Wernberg2011; @Wernberg2013; @Wernberg2016; @Mabin2013]. Given kelp’s preference for cold-temperate waters, ocean warming poses a serious threat to these ecosystems [@Wernberg2010; @Wernberg2011; @Verges2016]. In the absence of kelp, other types of systems may persist, for example algal turf, coralline reefs or urchin barrens [Verges2016]. These other systems support a different array of fauna and flora and are considered less diverse and less productive than kelp forests [@Dayton1985; @Connell2008; @Byrnes2011]. In future climate scenarios, species distributions are expected to shift poleward in response to changing ocean temperatures. Given kelps are cold-water species, changes in ocean temperatures, particularly warming, could have detrimental effects on their growth, survival, reproduction and recruitment.

Other abiotic drivers such as light [@Dayton1984; @Bolton1985; @Dieck1993; @Sjøtun2006; @Wernberg2008], sedimentation [@Duggins1990; @Vadas1992; @Schiel2006; Schiel2006] and water motion [@Hurd2000; @Kawamata2001; @Mudge2003; @Scrosati2004; @Wernberg2008; @Pedersen2012] have been shown to be important in the settlement, growth, and survival of kelp and biotic variables (e.g. see @Black1976; @Kennelly1983; @Dayton1984; @Harris1984; @Carpenter1990; @Sivertsen2006; @Harley2012) play an equally important role in structuring kelp communities, through alteration in top-down and bottom-up control mechanisms [@Breen1976; @Duggins1980; @Harris1984; @Estes1995; @Shears2002; @Arkema2009; @Blamey2012].

@@

## Statisical analyses

@@

### Community analysis

@@

flora_df <- read.csv(file = "data/flora.csv", check.names = FALSE)
fauna_df <- read.csv(file = "data/fauna.csv", check.names = FALSE)
benthic <- read.csv(file = "data/benthic.csv", check.names = FALSE)
cover_df <- read.csv(file = "data/cover.csv", check.names = FALSE)
kelp_df <- read.csv(file = "data/kelp.csv", check.names = FALSE)

@@

### Species composition

@@

plot_flora <- ggplot(data = flora_lng, aes(x = species, y= biomass, fill = site)) + 
              geom_bar(stat = "identity", show.legend = FALSE, position=position_dodge()) + 
              theme(axis.text.x=element_text(angle = 90, hjust = 1, vjust = 0.3)) +
              theme(panel.background = element_rect(fill = 'white', colour = 'white')) +
              ylab("") + 
              xlab("") +
              coord_flip()

@@

```{r biomass barplot, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap='Comparison of raw biomass species data between sites. Graph A represents flora and graph B represents fauna.'}
ggarrange(plot_flora, plot_fauna, ncol = 2, nrow = 1, labels = "AUTO", legend = "bottom", common.legend = TRUE, align = "none", heights = c(1.5,1))
```

## MDS

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

flora_mds <-  read.csv(file = "data/flora.csv", row.names = 'id', sep = ",", header = TRUE) 
site  <- flora_mds$site
flora_mds <- subset(flora_mds,select=-c(site))

flora_log <- decostand(flora_mds, method='log') # Log-transform data, accounts for possible over-estimate of species with a large biomass but low abundance
dist1<-vegdist(flora_log, method='bray')
nmds1 <- metaMDSiter(dist1,k=2)

nmds1_plot <- plot(nmds1, type = "p")

fauna_mds <-  read.csv(file = "data/fauna.csv", row.names = 'id', sep = ",", header = TRUE) 
site  <- fauna_mds$site
fauna_mds <- subset(fauna_mds,select=-c(site))
fauna_mds <- fauna_mds[which(rowSums(fauna_mds) > 0),]

fauna_log <- decostand(fauna_mds, method='range', na.rm = TRUE) # Log-transform data, accounts for possible over-estimate of species with a large biomass but low abundance
dist2<-vegdist(fauna_mds, method='bray')
nmds2 <- metaMDS(dist2, k = 3)

nmds2_plot <- plot(nmds2, type = "p")

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap= 'Multidimensional scaling of flora and fauna communities respectively.'}
ggarrange(nmds1_plot, nmds2_plot, ncol = 2, nrow = 1, labels = "AUTO", align = "none")
```
